---
title: "Finalize Phyloseq Object"
output: html_notebook
---

* Incorporate phylogenetic tree into phyloseq object
* Save master phyloseq objects

```{r}
library(dada2)
library(tidyr)
library(dplyr)
library(phyloseq)
library(ggplot2)
library(ape)
```

```{r}
DataDir = '/project/mmicrobe/Boyd/16S/DADA2files'
phyloseq_file = 'Boyd_phyloseq_thresh.rds'
tree_file = '/project/mmicrobe/Boyd/Fasttree/Master_wSulfo.tree'
```

```{r}
ps.thresh = readRDS(file.path(DataDir, phyloseq_file))
ps.thresh
```
```{r}
thresh.tree = read_tree(tree_file)


thresh.tree = root(thresh.tree, outgroup = 'X90478', r = TRUE)
is.rooted(thresh.tree)


```

# Create dummy phyloseq object to incorporate sulfolobus 
```{r}
ps0 <- prune_taxa(taxa_names(ps.thresh)[1:5], ps.thresh)
ps0 <- prune_samples(sample_names(ps0)[1:5], ps0)

otu_mat = matrix(1, nrow = 1, ncol = 1)
colnames(otu_mat) = "X90478"
rownames(otu_mat) <- "Root"
OTU <- otu_table(otu_mat, taxa_are_rows = FALSE)


SAM <- sample_data(ps.thresh)[1,]
SAM[,] <- NA
sample_names(SAM) <- "Root"
# or : rownames(SAM) <- "Reference"


tmp = tax_table(ps0) %>% as.data.frame()

tmp["X90478",] = c("Archaea", "Crenarchaeota", "Thermoprotei", "Sulfolobales", "Sulfolobaceae","Saccharolobus" )
tmp = tmp %>% as.matrix()

tmp = tmp["X90478",, drop = FALSE]



rootTax = tax_table(tmp)
rootTax
ps.root <- phyloseq(OTU, SAM, rootTax)
ps.root %>% taxa_names()

ps.merged =  merge_phyloseq(ps.thresh, ps.root)
ps.merged

tax_table(ps.merged) %>% tail()
otu_table(ps.merged)[1:5,7529]
```


# Join bacterial tree with phyloseq
* This will result in the loss of archaea so need to use both phyloseq objects
```{r}
ps = merge_phyloseq(ps.merged, thresh.tree)

#force polytomies to dichotomies
phy_tree(ps) = multi2di(phy_tree(ps), random = FALSE)
ps


```
```{r}
#remove mitochondria
ps = subset_taxa(ps, Family != "Mitochondria" | is.na(Family))
ps = subset_taxa(ps, Order != "Chloroplast" | is.na(Order))
ps
```
```{r}
saveRDS(ps, file.path(DataDir,'Boyd_phyloseq_thresh_wTree.rds'))
```

# Check duplicates
```{r}
sd = ps %>% sample_data() %>% as.data.frame()
sd %>% head()
Dups = sd %>% 
  group_by(SampleID) %>%
  summarize(n = length(SampleID)) %>%
  filter(n > 1)
Dups

Dups = Dups$SampleID %>% as.character() 
Dups
```

```{r}
ps.s = ps %>% subset_samples(!is.na(TimePoint))
ordu = ordinate(ps.s, "PCoA", "bray", weighted=TRUE)
p = plot_ordination(ps.s, ordu, justDF = TRUE)
plot_ordination(ps.s, ordu)
```

```{r}
#setwd(figDir)
#cs = brewer.pal(3,"Dark2")
p = p %>%
  mutate(Dup = ifelse(SampleID %in% Dups[16], "Duplicate", "Single"))
p1 = ggplot(p, aes(x = Axis.1, y = Axis.2)) +
geom_point(aes(color = Dup),  size = 3, stroke = 1.5) +
#theme_pub() +
#scale_shape_manual(values=c(22,21,23), guide = guide_legend(title = "Soil", ncol = 1))+
#scale_color_manual(values=c('Orange','Black'))+
scale_color_manual(values = c('Orange','Black'))+
labs(x = "PCoA 1 [44.3%]", y = "PCoA 2 [17.3%]") +
ggtitle("PCoA of Weighted UniFrac Distances")+
facet_wrap(~TimePoint, ncol = 2)

#guides(fill=guide_legend(override.aes=list(colour= cs), ncol = 3)) + 
#theme(legend.position="bottom") 

#color = "grey20",

p1


```

* Duplicates do not cluster well.  Show greater separation than in any dataset I have worked with.  Possible that there is less overall variation compared to other studies.  Plot treatment differences to see if there is any separation

```{r}
p %>% head()
#p = p %>%
#  mutate(Dup = ifelse(SampleID %in% Dups[16], "Duplicate", "Single"))
#  filter(!SampleID %in% Dups)
ptmp = p %>% filter(Treatment %in% c("Silage-NCC", "Corn-RCC"),Soil_depth == 1)
ptmp
p1 = ggplot(ptmp, aes(x = Axis.1, y = Axis.2)) +
geom_point(aes(fill = Treatment),  size = 2, shape = 22) +
#theme_pub() +
#scale_shape_manual(values=c(22))+#, guide = guide_legend(title = "Soil", ncol = 1))+
#scale_color_manual(values=c('Orange','Black'))+
scale_fill_manual(values = c('Orange','Black'))+
labs(x = "PCoA 1 [44.3%]", y = "PCoA 2 [17.3%]") +
ggtitle("PCoA of Weighted UniFrac Distances")+
facet_wrap(~TimePoint, ncol = 2)

#guides(fill=guide_legend(override.aes=list(colour= cs), ncol = 3)) + 
#theme(legend.position="bottom") 

#color = "grey20",

p1
```
* Some structure, but clearly the most extreme treatments are not fully separating  And some samples are very, very similar.  Bring to PC and explore control samples and mock community.
